EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-05-20 05_examples.tex
Intermediate commit. Adding BHZ nonSU2

@@

\subsection{Multi-orbital impurity with Hubbard-Kanamori interaction (TRIQS)}
\label{SecExamplesTRIQS}

@@

@@

\subsection{Bethe lattice DMFT (w2dynamics)}\label{SecExamplesBetheDMFTw2d}
Let us revisit the example of Sec.~\ref{SecExamplesBetheDMFT} to
demonstrate usage of the w2dynamics interface how to perform such a
calculation instead. Unlike when the Fortran API is used, none of the
five steps of the iterative algorithm listed there are left to the
user to implement themselves. Instead, in addition to possibly
non-default values for \NAME input variables, the model specification
must also be added to a w2dynamics parameter file and depending on the
model, further input files might need to be provided. Further
differences are that initialization is performed in w2dynamics by
starting from a self-energy of zero or reading a self-energy value
from an old file, which results in the initial Weiss field that the
initial bath parameters are fitted to, and that automatic convergence
checking is not currently available. Finally, using the same settings
as for the Fortran example we arrive at the w2dynamics parameter file
shown in the following listing.

\subsection{Interacting Bernevig-Hughes-Zhang (Fortran)}

>>>>

\subsubsection{W2Dynamics solution}\label{SecExamplesBetheDMFTw2d}
To demonstrate use of the w2dynamics interface, here we briefly
discuss how to solve the DMFT problem for the Bethe lattice using
W2Dynamics interface. \fixme{For definitiness we compare the finite low
temperature results obtained via the two solvers: ED and CTQMC.  }

Unlike when the Fortran API is used, none of the
five steps of the iterative algorithm listed there are left to the
user to implement themselves. Instead, in addition to possibly
non-default values for \NAME input variables, the model specification
must also be added to a w2dynamics parameter file and depending on the
model, further input files might need to be provided. Further
differences are that initialization is performed in w2dynamics by
starting from a self-energy of zero or reading a self-energy value
from an old file, which results in the initial Weiss field that the
initial bath parameters are fitted to, and that automatic convergence
checking is not currently available. Finally, using the same settings
as for the Fortran example we arrive at the w2dynamics parameter file
shown in the following listing.

@@

\subsection{Multi-orbital impurity with Hubbard-Kanamori interaction}
\label{SecExamplesTRIQS}

@@

\fixme{Add Hamiltonian?}
\fixme{}

@@

\subsection{Interacting Bernevig-Hughes-Zhang model}
In this section we focus on the DMFT solution of the interacting
Bernevig-Hughes-Zhang (BHZ).
As shown in Ref.~\onlinecite{AmaricciPRB2022}, at moderate
interactions this model admits an excitonic phase in which opposite
orbitals electrons and holes across the band gap bound into a coherent phase.

We consider a system of two-orbital electrons on a square
lattice interacting through a Hubbard-Kanamori term and describing
a quantum spin Hall insulator.
We consider a suitable matrix basis in terms of the Dirac
matrices $\Gamma_{a\a}=\sigma_a\otimes \tau_\a$, where $\sigma_a$ and
$\tau_\a$ are Pauli matrices, respectively, in the spin and orbital
pseudo-spin space. The  model Hamiltonian reads
$$
H = \sum_{k}\psi_{k}^\dagger H(k)\psi_{k} + H_{\rm int}
$$
where $\psi_{k}=[c_{1\uparrow k}, c_{2\uparrow k},
c_{1\downarrow k}, c_{2\downarrow k} ]$ is the spinor collecting
annihilation operator $c_{a\sigma k}$ destructing an electrons at
orbital $a=1,2$ with spin  $\sigma=\up,\dw$ and lattice momentum
$k$. The non-interacting part of the Hamiltonian is:
$$
H(k) = \left[M-2t(\cos{k_x}+\cos{k_y}) \right]\Gamma_{03} +
   \lambda\sin{k_x}\Gamma_{31} -   \lambda\sin{k_y}\Gamma_{02}
$$
where $M$ is the mass term, which plays the role of a crystal
field splitting among the orbitals. The presence of this term breaks
the symmetry in the orbital pseudo-spin channel.
The  interaction reads: 
$$
   H_{\rm int} = (U-J)\frac{\hat{N}(\hat{N}-1)}{2} - J\left( \frac{1}{4}\hat{N}^2 +
   \hat{S_z}^2 - 2 \hat{T_z}^2\right)
 $$
 where $\hat{N}=\tfrac{1}{2}\psi_i^\dagger \Gamma_{00}\psi_i$ is the
total density operator,
$\hat{S_z}=\tfrac{1}{2}\psi_i^\dagger \Gamma_{30}\psi_i$ is the total
spin polarization operator and $\hat{T_z}=\tfrac{1}{2}\psi_i^\dagger
\Gamma_{03}\psi_i$ is the orbital pseudo-spin polarization operator.
This form corresponds to the density-density part of the
Kanamori interaction. We neglect the pair-hopping and spin-flip purely
for numerical reasons.

In the non-interacting regime, i.e. $U=J=0$ this model describes a
quantum spin Hall insulator (QSHI) for $M<4t$ and a trivial Band
Insulator (BI) for $M>4t$.
The transition point at $M=4t$ describes the formation of a gapless
Dirac state with linear band crossing at $k=[0,0]$.

The code implementation follows the same guidelines discusses above
for the other examples. There are two specific modifications requied
to the  user. The first is to generate the Hamiltonian $H(k)$ on a discretized Brillouin zone in
order to obtain both the local non-intearcing Hamiltonian
$H_{loc}=\tfrac{1}{N_k}\sum_k H(k)$ and get the local interacting
Green's function
$$
G_{loc}(z) = \tfrac{1}{N_k}\sum_k \left[z+\mu-H(k)-\Sigma(z) \right]^{-1}
$$
where $\z\in\CCC$, $\Sigma(z)$ is the self-energy function obtained from the DMFT
solution of the problem and $N_k$ is the number of $k$-points. 
The second is to  construct the ``renormalized'' topological
Hamiltonian
$$
H_{top} = \sqrt{Z}[H(k) + \Re\Sigma(\omega\to0)]\sqrt{Z}, 
$$
where $Z$ is the renormalization constant. $H_{top}$ describes the
low-energy properties of many-body problem and can be used to obtain a 
topological characterization of the interacting system.

\subsubsection{Results}
To capture the possible symmetry breaking in any excitonic channel
driven by the local interaction, we consider the vector order
parameter
$\vec{E}=[E_0,E_x,E_y,E_z]$ where $E_a=\langle\psi^\dagger
\Gamma_{a1}\psi \rangle$.
The first component $E_0$ describes the singlet excitonic state,
whereas the remaing ones correspond to the triplet states with
different spin orientation.   
The analysis of the strong-coupling regime as well as impurity
excitonic susceptibility available in \NAME suggest the possible
instability to an in-plane spin-triplet exciton phase, i.e. with $E_x$
and/or $E_y$ different from zero.
Interestingly, this state breakes several symmetries including
time-reversal and spin SU(2) which protect the topological state.

Thus, here we illustrate the DMFT solution of the excitonic phase as a way to
illustrate the ability of \NAME to capture matter phases with broken
spin-symmetry ({\tt ed\_mode=nonsu2}) and the use of {\tt
  bath\_type=replica}.
\begin{lstlisting}[style=fstyle,numbers=none,basicstyle={\scriptsize\ttfamily}]
   !> Set $H_{loc}$ in \NAME
   allocate(Hloc(Nso,Nso))
   Hloc = sum(Hk,dim=3)/Lk
   where(abs(dreal(Hloc))<1d-6)Hloc=zero
   call ed_set_hloc(Hloc)
   !> Setup the replica bath basis using  $M$, $E_0$, $E_z$ and $E_x$ as symmetries
   allocate(lambdasym_vector(Nbath,4))
   allocate(Hsym_basis(Nso,Nso,4))
   Hsym_basis(:,:,1)=Gamma03  ;lambdasym_vector(:,1)= Mh
   Hsym_basis(:,:,2)=Gamma01 ;lambdasym_vector(:,2)= sb_field
   Hsym_basis(:,:,3)=Gamma31 ;lambdasym_vector(:,3)= sb_field
   Hsym_basis(:,:,4)=Gamma11 ;lambdasym_vector(:,4)=-sb_field
   !> Set the replica bath
   call ed_set_Hreplica(Hsym_basis,lambdasym_vector)
   !> Get bath dimension and allocate user bath to this size
   Nb=ed_get_bath_dimension(4)   !There are 4 symmetries
   allocate(Bath(Nb))
   !
   !> Initialize the ED solver (bath is guessed or read from file) 
   call ed_init_solver(bath)
\end{lstlisting}

In the following we consider the case of $M=1$ which corresponds to a
QSHI in the non-interacting limit and $J/U=0.25$. 
%
The main effect of the interaction is enclosed in the renormalized
mass term $M_{eff}=M+\tfrac{1}{4}\Tr{\Re\Sigma(i\omega\to0)}$. The
real-part of the self-energy being proportional to $\Gamma_{03}$
corrects the mass term with respect to its bare value. To leading
order this correction is given proportional to $\tfrac{U-5J}{4}\langle
\hat{T}_z\rangle$. So, for our choice of parameters, the effect of
interaction would be to reduce the (effective) mass term and thus
reducing the topological band gap. In the strong-coupling limit the
two orbitals get populated with one electron per site, reaching the
conditions for the formation of a Mott insulating state.
In between, at inter-mediate coupling, we observe the formation of a
region of exciton condensation.
In panel (A) of \figu{figEx4} we report the evolution of the order
parameter

\begin{figure}[t!]
  \includegraphics[width=\linewidth]{figures/figBHZ.pdf}
    \caption{\label{figEx4}%
      \textbf{Topological and Exciton Transition.}
      (a)
    }
\end{figure}
